The Dynamics of Anthropomorphic Effects in Human-Robot Interaction

### 2014-03-25 paper.tex
Completed section 'A few result from the field' in intro
Still wondering if the referenced figures should not be moved there as well

To support the reflection, it may be useful to consider a few experimental
results. We observed anthropomorphism in several short- and long-term human-robot
interaction studies that we carried out with embodied robots, mostly in
real-world settings. There is not much research that investigates
anthropomorphism based on real interaction between a human and a robot, and
above all in natural settings and over extended periods of time.

Figure~\ref{fig:roomba-perception} gives a first perspective on how naive
users perceive robots over a long period of time. Here, \emph{Roomba} vacuum cleaning robots
were given to nine households for a 6-months period. The aim was to explore how
people use and perceive the robot over time, and how they generally live with
the Roomba in their home. At various time points, users were asked to rate
their perception of different features of the robot on a 7-point rating scale. As
Figure~\ref{fig:roomba-perception} shows, some features remained stable over time
(\emph{ease of use} and \emph{fun}), and others decreased: the perceived
\emph{usefulness}, \emph{intelligence} and \emph{impact} on their household.
\emph{intelligence} and \emph{impact} reflect the cognitive engagement of the
user towards the machine: we can assume that one does not care to engage in a
(cognitive) interaction with an artifact perceived as non-intelligent and with
no impact on his/her life. Even if one can argue that the Roomba is not designed
to foster interaction, the question is raised: how to sustain engagement between
a robot and a human over a long period of time? And, as a pre-requisite, do we
actually understand the psychological phenomenon, and can we effectively account
for the dynamics of these interactions?

Those questions are certainly not new, but paradoxically, while anthropomorphism
in robotics is a commonly discussed and studied trait of human-robot
interaction, it appears that it lacks a formal ground and unified understanding to
take into account the wide range of phenomena that it encompasses.

>>>>

To support the reflection, it may be useful to consider a few experimental
results. We observed anthropomorphism in several short- and long-term human-robot
interaction studies that we carried out with embodied robots, both in
ecologically valid real-world environments and controlled lab settings. Those
experiments are presented in details in section~\ref{sec:overview-studies}. We
only highlight here some qualitative results in order to trigger some initial
thoughts.

Figures~\ref{fig:roomba-activities} and~\ref{fig:roomba-perception} (on
page~\pageref{fig:roomba-perception}) give a first perspective on how naive
users interact and perceive robots over a long period of time. Here,
\emph{Roomba} vacuum cleaning robots were given to nine households for a
6-months period. The aim was to explore how people use and perceive the robot
over time, and how they generally live with the Roomba in their home. At various
time points, users were asked to rate their perception of different features of
the robot on a 7-point rating scale. As Figure~\ref{fig:roomba-perception}
shows, some features remained stable over time (\emph{ease of use} and
\emph{fun}), and others decreased: the perceived \emph{usefulness},
\emph{intelligence} and \emph{impact} on their household.  \emph{intelligence}
and \emph{impact} reflect the cognitive engagement of the user towards the
machine: we can assume that one does not care to engage in a (cognitive)
interaction with an artifact perceived as non-intelligent and with no impact on
his/her life (and the level of activity pictured on
Figure~\ref{fig:roomba-activities} qualitatively confirms this intuition). Even
if one can argue that the Roomba is not designed to foster interaction, it
raises a first question: how to sustain engagement between a robot and a human
over a long period of time? And, as a pre-requisite to answer, do we actually
understand the psychological phenomenon, and can we effectively account for the
dynamics of these interactions?

Those questions are certainly not new, but paradoxically, while anthropomorphism
in robotics is a commonly discussed trait of human-robot interaction, it appears
that we lack formal grounds and models that would account for the long-term
dynamics of these interactions.

Figure~\ref{fig:qualitative-score} (page~\pageref{fig:qualitative-score}) calls
for other initial remarks. This diagram comes from a study where 13 pairs of
children (4-5 years old) played for about 20 minutes a Domino game with a robot
(the robot was tasked to carry domino pieces from one child to the other). The
$x$-axis represents the number of social interactions between the children and
the robot (like talking or showing objects), based on video annotations. On the
$y$-axis, we have a synthetic qualitative measurement of anthropomorphic
projections based on two questionnaires. The correlation between both these
values is significantly \emph{negative}: the more the children engaged into
interaction, the less prone to anthropomorphic projections they were. This
diagram also highlights an interesting cluster of children who did not interact
much with the robot, and still reported a strong anthropomorphic perception of
the robot.

These results may seem counter-intuitive at first sight, and illustrate that the
mechanisms that relate anthropomorphism to adoption and engagement are
non-trivial. In section~\ref{sec:cognition-neuroscience}, we sketch a model of
the cognitive correlates of anthropomorphism that partially account for these
results.
